---
title: "Case study 4 (guided online only)"
author: "Owen"
date: "2/10/2019"
output: html_document
---

General article about dietary restriction (DR): https://www.nia.nih.gov/health/calorie-restriction-and-fasting-diets-what-do-we-know

Evolution under dietary restriction increases male reproductive performance without survival cost

Zajitschek F, Zajitschek SRK, Canton C, Georgolopoulos G, Friberg U, Maklakov AA

Date Published: January 29, 2016

Article: https://doi.org/10.1098/rspb.2015.2726

DOI: https://doi.org/10.5061/dryad.2fp25

"We populated each cage with 150 males and 150 females, and set up four replicate cages per diet. We used three diets that differed in their yeast concentration (40, 100 or 270 g per 1 litre of diet: from here on DR, standard and high protein diet, respectively""

"108 experimental vials."

"Flies were provided with one of the three diets used in the experimental evolution cages, with three replicate vials per cage and diet combination (12 cages, three diets per cage, four replicates)."

```{r message=FALSE, echo=TRUE, warning=FALSE}
## Load the libraries we use
library(dplyr)
library(ggplot2)
library(readr)
library(stringr)
library(lubridate)
library(skimr)
library(tidyr)
library(ggbeeswarm)
library(taxize)
```


```{r}
survival <- read_csv("www/dryad Zajitschek et al 2016 ProcB Data male survival.csv")
```


```{r}
repr_perf <- read_csv("www/dryad Zajitschek et al 2016 ProcB Data male repro perf.csv")
```

## Checklist

### Before working in R
question
hypothesis
study methods and design
numbers of "things"
type of response variable
variables and their meaning of variables (given as comments in the Excel file)

### After importing the data into R
number of rows and columns
tidy?
variable types
duplicates
NAs
fix dates
replace codes with informative words
variable entries, e.g. levels of characters, ranges of numerics
numbers of "things"

## Calculate response variable(s) (if required)


## Inspect shapes (distributions)


## Inspect relationships


